---
title: "SeaFlow Example Preparation"
output: html_document
---

In this document, we create the `protoclust` object that will be used in the SeaFlow `protoshiny` example in the paper.

```{r}
library(protoclust)
library(tidyverse)
```

We begin by loading in the distance matrix.  Each entry of this matrix corresponds to the earth movers distance between a pair of cytograms. 

We thank Francois Ribalet and Mattias Cape for providing this distance matrix to us.  It was computed as part of the following poster:

> Cape, M. R., Ribalet, F., Bien, J., Hyun, S. & Armbrust, E. V. (2020), Ob14f-0437 - Determining Ecological Provinces from Optical Cytometric Data in the North Pacific Ocean, in ‘Ocean Sciences Meeting’, San Diego, CA.
URL: https://agu.confex.com/agu/osm20/meetingapp.cgi/Paper/657891

Because github limits the size of each file to be less than 100Mb, we have saved the upper triangle of the distance matrix as two vectors.  We load these and then reconstruct the symmetric distance matrix in what follows:

```{r}
wvec1 <- readRDS("wass_vec1.RDS")
wvec2 <- readRDS("wass_vec2.RDS")
wvec <- c(wvec1, wvec2)
n <- (1 + sqrt(1 + 8 * length(wvec))) / 2 # Find n based on n choose 2
stopifnot(choose(n, 2) == length(wvec))
dd <- matrix(0,n,n)
dd[upper.tri(dd)] <- wvec
dd <- dd + t(dd)
```

This creates a `r nrow(dd)`-by-`r ncol(dd)` distance matrix.  We also load the timestamps corresponding to these cytograms.

```{r}
date.vec <- readRDS("SCOPE_16_3min_dates.RDS")
```

Here's a look at the format of the timestamps:

```{r}
date.vec[1:2]
```

We will use these timestamps as labels; however, we shorten their format a bit for readability:

```{r}
times <- as.character(date.vec) %>% 
  str_remove("2016-") %>% 
  str_remove("\\+00:00$") %>% 
  str_replace(" ", "_h")
```

Let's add these as row and column labels to the distance matrix.

```{r}
rownames(dd) <- colnames(dd) <- times
```

Let's have a look at the top corner of the distance matrix:

```{r}
dd[1:3, 1:3]
```

Now we perform the hierarchical clustering.

```{r}
clustering_time <- system.time({hc <- protoclust(dd)})
clustering_time
```

For our `protoshiny`, we will want to have thumbnails for each observation as a different way of labeling observations other than timestamps.  To do this, we add a character vector `hc$img` of paths to the thumbnails.

```{r}
hc$img <- paste0("figs_", str_pad(1:n, 4, pad="0"),".png")
save(hc, file = "seaflow_hc.Rdata")
```

