---
title: "Movies Data Download and Preparation"
output:
  html_document:
    df_print: paged
---

# Download

We download the MovieLens 25M Dataset from here https://grouplens.org/datasets/movielens/25m/.  The paper to cite for this data is here: http://files.grouplens.org/papers/harper-tiis2015.pdf

```{r, eval = FALSE}
download.file("http://files.grouplens.org/datasets/movielens/ml-25m.zip",
              destfile = "ml-25m.zip")
unzip("ml-25m.zip")
```
Let's remove the files that pertain to actions of specific users since we will not need them here.  We'll be using the "genomes" which were learned from this user-level data.

```{r, eval = FALSE}
system("rm ml-25m/ratings.csv ml-25m/tags.csv")
```

# Look at data

```{r}
library(tidyverse)
gs <- read_csv("ml-25m/genome-scores.csv")
gt <- read_csv("ml-25m/genome-tags.csv")
mo <- read_csv("ml-25m/movies.csv")
```

In this paper, http://files.grouplens.org/papers/tag_genome.pdf, the authors describe the Tag Genome, a vector space embedding of movies.  Each dimension of this vector space gives the relevance score of this movie to a particular tag.  Tagging of movies is done by users, but the idea of the tag genome is to get continuous relevance scores from this raw tagging data.

The data frame `gs` gives for each movie-tag pair, how relevant that movie is to that tag.  For example, here are the five most relevant tags for the movie Groundhog Day: 
```{r}
groundhog_id <- mo %>% 
  filter(str_detect(title, "Groundhog")) %>% 
  pull(movieId)
gs %>% 
  filter(movieId == groundhog_id) %>% 
  top_n(5, relevance) %>% 
  arrange(desc(relevance)) %>% 
  left_join(gt, by = "tagId") %>% 
  select(tag, relevance)
```

# Prepare

As expected, every movie has the same number of tag relevance scores:
```{r}
gs %>% 
  count(movieId) %>% 
  pull(n) %>% 
  range()
```

Let's make this into a dense matrix, with movies as rows and tags as columns:
```{r}
mid <- unique(gs$movieId)
tid <- unique(gs$tagId)
stopifnot(gs$movieId == rep(mid, each = length(tid)))
stopifnot(gs$tagId == rep(tid, times = length(mid)))
x <- matrix(gs$relevance, nrow = length(mid), byrow = TRUE)
rownames(x) <- mid
colnames(x) <- tid
movie_info <- tibble(movieId = mid) %>% 
  inner_join(mo, by = "movieId")
stopifnot(length(mid) == nrow(movie_info))
rownames(x) <- movie_info$title
tag_info <- tibble(tagId = tid) %>% 
  inner_join(gt, by = "tagId")
stopifnot(tid == tag_info$tagId)
colnames(x) <- tag_info$tag
```

There are `r nrow(x)` movies and `r ncol(x)` tags.

Let's check that we did this correctly:
```{r}
x[rownames(x) %>% str_which("Groundhog"),] %>% sort(decreasing = TRUE) %>% .[1:5]
x[rownames(x) %>% str_which("Willy Wonka"),] %>% sort(decreasing = TRUE) %>% .[1:5]
```
Looks like we did!

# Protoclust

Let's compute the pairwise correlations between all movies:

```{r}
system.time({movie_cor <- cor(t(x))})
d <- as.dist(1 - movie_cor)
rm(movie_cor)
```

And now let's run protoclust:

```{r}
library(protoclust)
system.time({hc <- protoclust(d)})
save(hc, movie_info, file = "movies_hc.Rdata")
```

